---
title: "6_24"
output: html_notebook
---

- try row_km to split the dendrogram -> check performance of clusters

# Load packages

```{r}
library(readr)
library(readxl)
library(DunedinPACE)
library(data.table)
library(pheatmap)
library(ComplexHeatmap)
library(circlize)
library(dplyr)
library(tidyverse)
library(splines)
library(clusterProfiler)
library(org.Hs.eg.db)
library(ReactomePA)
```

# Extract CpG sites from Investigated Clocks

```{r}
Hannum_CpGs <- read_excel("CpGsToInvestigate/hannum_cpgs.xlsx")
Hannum_CpGs <- Hannum_CpGs$Marker #71
Levine_CpGs <- read.csv("CpGsToInvestigate/levine_cpgs.csv", stringsAsFactors=FALSE)
Levine_CpGs <- Levine_CpGs[-1, ]
Levine_CpGs <- Levine_CpGs$CpG #513
Horvath_CpGs <- read.csv("CpGsToInvestigate/horvath_cpgs.csv", stringsAsFactors=FALSE)
Horvath_CpGs <- Horvath_CpGs[-(1:3), 1, drop=FALSE]
Horvath_CpGs <- Horvath_CpGs[, 1] #353
McEwen_CpGs <- read.csv("CpGsToInvestigate/mcewen_cpgs.csv")
McEwen_CpGs <- McEwen_CpGs$CPG #94
Wu_CpGs <- read_excel("CpGsToInvestigate/aging-11-102399-s003..xlsx")
Wu_CpGs <- Wu_CpGs[-1, ]
Wu_CpGs <- Wu_CpGs$CpGs #111
Belsky_CpGs <- getRequiredProbes(backgroundList=FALSE)
Belsky_CpGs <- unlist(Belsky_CpGs) #173
load("CpGsToInvestigate/epitoc.Rd")
Teschendorff_CpGs <- as.data.frame(dataETOC2.l[1])
Teschendorff_CpGs <- rownames(Teschendorff_CpGs) #163
Shireby_CpGs <- readLines("CpGsToInvestigate/CorticalClockCoefs.txt")[-1]
Shireby_CpGs <- sapply(strsplit(Shireby_CpGs, " "), `[`, 1)
Shireby_CpGs <- unlist(Shireby_CpGs) #347
Lu2_CpGs <- read_excel("CpGsToInvestigate/lu2.xlsx")
Lu2_CpGs <- rbind(colnames(Lu2_CpGs), Lu2_CpGs)
Lu2_CpGs <- Lu2_CpGs[, 1]
Lu2_CpGs <- as.character(Lu2_CpGs[[1]]) #140
clock_cpgs <- list(
  McEwen_CpGs = McEwen_CpGs,
  Wu_CpGs = Wu_CpGs,
  Hannum_CpGs = Hannum_CpGs,
  Horvath_CpGs = Horvath_CpGs,
  Levine_CpGs = Levine_CpGs,
  Belsky_CpGs = Belsky_CpGs,
  Teschendorff_CpGs = Teschendorff_CpGs,
  Shireby_CpGs = Shireby_CpGs,
  Lu2_CpGs = Lu2_CpGs
)

cpgs <- unlist(clock_cpgs)
cpgs <- unique(cpgs) #1868 total sites

print(paste0("Total Unique CpG sites from 9 clocks explored: ", length(cpgs)))

#write.csv(cpgs, "7_3_overlap.csv")
```
To test

```{r}
generateMatrix <- function(gender_state=1, clock_state=TRUE, output_file_path1, output_file_path2) {
  # Initialize input path
  list_files <- list.files(path="/Users/kchen/OneDrive/Documents/methylation/combined_split3")
  # Extract number of splits
  num_files <- length(list_files)
  
  # initial_cpg_list contains the full list of CpGs (from preinitializing chunk)
  # NUMCPGSITE is length of initial_cpg_list (from preinitializing chunk)
  
  # initial_clock_list contains the list of Clock CpGs
  # NUMCPGSITECLOCK is length of initial_clock_list
  
  # Create empty output matrix
  if (clock_state) {
    B <- createEmptyMatrix(NUMCPGSITECLOCK, initial_clock_list, 0, 100)
    S <- createEmptyMatrix(NUMCPGSITECLOCK, initial_clock_list, 0, 100)
  } else {
    B <- createEmptyMatrix(NUMCPGSITE, initial_cpg_list, 0, 100)
    S <- createEmptyMatrix(NUMCPGSITE, initial_cpg_list, 0, 100)
  }
  
  # Loop through all splits
  for (i in 1:num_files) {
    # Read split from folder
    temp <- fread(paste0("/Users/kchen/OneDrive/Documents/methylation/combined_split3/", list_files[i]))
    # Format:
    # - First two columns are identical: "V1", "cpg"
    #  - contain "age" then CpG identifiers
    # - Column names are GSM IDs
    # - First Row contains ages
    
    # Format split
    temp <- as.data.frame(temp)
    # Set rownames
    rownames(temp) <- temp[, 1]
    temp <- temp[, -c(1, 2)]
    # Round non-integer ages
    temp["age", ] <- round(temp["age", ])
  
    # Filter for desired samples (gender-stratified)
    if (gender_state == 1) {
      temp <- temp[, colnames(temp) %in% all_samples]
    } else if (gender_state == 2) {
      temp <- temp[, colnames(temp) %in% male_samples]
    } else if (gender_state == 3) {
      temp <- temp[, colnames(temp) %in% female_samples]
    } else {
      print("Error")
      break
    }
    
    # Reshape temp for data.table operations
    temp_melt <- melt(temp, id.vars = "age")
    setDT(temp_melt)
    
    # Calculate mean and sd for each age group
    results <- temp_melt[, .(mean_value = mean(value, na.rm = TRUE), sd_value = sd(value, na.rm = TRUE)), by = .(age, variable)]
    
    # Reshape results back to wide format
    mean_wide <- dcast(results, variable ~ age, value.var = "mean_value")
    sd_wide <- dcast(results, variable ~ age, value.var = "sd_value")
    
    # Assign back to matrices
    for (j in 0:100) {
      age_col <- paste0("age_", j)
      B[ident, age_col] <- mean_wide[[as.character(j)]]
      S[ident, age_col] <- sd_wide[[as.character(j)]]
    }
    
    # Logging
    print(paste0("Processing Split ", i))
    gc()
  }
  
  print("Writing files now")
  fwrite(B, output_file_path1, row.names = TRUE)
  fwrite(S, output_file_path2, row.names = TRUE)
}

generateMatrix(gender_state = 1, clock_state = FALSE, "7_19_all_matrix_avg.csv", "7_19_all_matrix_sd.csv")

```

```{r}

# Filter for avg_beta correlation > corr_cutoff, avg_beta abs change > abs_change_cutoff,
# less than num_sd_cutoff ages have SD > sd_coeff_cutoff * sqrt(avg_beta * (1-avg_beta))
# gender_state: 1 = no stratification, 2 = male, 3 = female
filterCpGResults <- function(corr_cutoff, abs_change_cutoff, sd_coeff_cutoff, num_sd_cutoff, gender_state) {
  # Initialize input path
  list_files <- list.files(path="/Users/kchen/OneDrive/Documents/methylation/combined_split3")
  # Extract number of splits
  num_files <- length(list_files)
  # Create results format (Key: Age Window, Value: All sites fulfilling criteria)
  results <- data.frame()
  
  # initial_cpg_list contains the full list of CpGs (from preinitializing chunk)
  # NUMCPGSITE is length of initial_cpg_list (from preinitializing chunk)
  
  # Loop through every split
  for (i in 1:num_files) {
    # Read split from folder
    temp <- fread(paste0("/Users/kchen/OneDrive/Documents/methylation/combined_split3/", list_files[i]))
    # Format:
    # - First two columns are identical: "V1", "cpg"
    #  - contain "age" then CpG identifiers
    # - Column names are GSM IDs
    # - First Row contains ages
    
    # Format split
    temp <- as.data.frame(temp)
    rownames(temp) <- temp[, 1]
    temp <- temp[, -c(1, 2)]
    # Round non-integer ages
    temp["age", ] <- round(temp["age", ])
  
    # Filter for desired samples (gender-stratified)
    if (gender_state==1) {
      temp <- temp[, colnames(temp) %in% all_samples]
    } else if (gender_state==2) {
      temp <- temp[, colnames(temp) %in% male_samples]
    } else if (gender_state==3) {
      temp <- temp[, colnames(temp) %in% female_samples]
    } else {
      print("Error")
      break
    }
    
    # Save identifiers to examine
    ident <- rownames(temp[-1, ])
    NUMCPGSITE <- length(ident)
    
    # Loop through age windows
    for (start_age in seq(0, 60, by=5)) {
      end_age <- start_age + 20
      
      # Create Empty Average Matrix
      A <- createEmptyMatrix(NUMCPGSITE, ident, start_age, end_age)
      # Create Empty SD Matrix
      S <- createEmptyMatrix(NUMCPGSITE, ident, start_age, end_age)
      
      # Filter for only samples in age range [0, 20], [20, 40]...
      # Ages 20, 40, 60 will overlap between windows
      des_cols <- which(temp["age", ]>= start_age & temp["age", ]<=end_age)
      temp2 <- temp[, des_cols]
      
      # Loop through every age in age window
      for (j in start_age:end_age) {
        des_cols2 <- which(temp2["age", ]==j)
        # Ensure samples are present at age j
        if (length(des_cols2)>0) {
          temp3 <- temp2[, des_cols2]
          # Remove age row
          temp3 <- temp3[-1, ]
          setDT(temp3)
          
          #Vectorize computations for all sites in split i for age j
          average_vector <- temp3[, .(final = rowMeans(.SD, na.rm=TRUE)), by=.I]
          average_vector <- as.data.frame(average_vector)
          sd_vector <- temp3[, .(final = apply(.SD, 1, sd, na.rm=TRUE)), by=.I]
          sd_vector <- as.data.frame(sd_vector)
          
          # Assign back to matrices
          A[, paste0("age_", j)] <- average_vector$final
          S[, paste0("age_", j)] <- sd_vector$final
          
          # Logging
          #print("Processing age ", j)
        }
      }
      
      # Both matrices for age_window have finished computing
      # Start filtering process
      
      # Initialize results matrix
      results <- data.frame(matrix(nrow=nrow(A), ncol=3))
      rownames(results) <- rownames(A)
      colnames(results) <- c("sd_res", "cor_res", "abs_change_res")
      
      # Fill in sd_res
      for (k in 1:nrow(A)) {
        count <- 0
        # Count how many are above SD threshold
        for (l in 1:ncol(A)) {
          adj_avg <- sd_coeff_cutoff * sqrt(A[k, l] * (1-A[k, l]))
          sd <- S[k, l]
          if (!is.na(adj_avg) && !is.na(sd) && (sd > adj_avg)) {
            count <- count + 1
          }
        }
        results[k, "sd_res"] = count
      }
      
      # Fill in cor_res
      results$cor_res <- apply(A, 1, function(x) cor(x, start_age:end_age))
      
      # Fill in abs_change_res
      results$abs_change_res <- apply(A, 1, function(x) max(x) - min(x))
      
      # Filter for sites that fit criterion
      filtered_results <- results[results$sd_res<num_sd_cutoff & abs(results$cor_res)>corr_cutoff & results$abs_change_res>abs_change_cutoff, ]
      selected_sites <- rownames(filtered_results)
      
      # Add to running list (from previous splits)
      results_by_window[[paste(start_age, "to", end_age)]] <- c(results_by_window[[paste(start_age, "to", end_age)]], selected_sites)
    }
    gc()
  }
  return(results)
}

visualizeResults <- function(results) {
  for (i in 1:length(results)) {
    print(paste0("Start age: ", (i-1)*5, " End age: ", (i-1)*5+20, " Filtered: ", ... = length(results[[i]])))
  }
  all_sites <- unlist(results)
  unique_sites <- unique(all_sites)
  print(length(unique_sites))
}

results1 <- filterCpGs(0.4, 0.2, 0.5, 3, 1)
visualizeResults(results1)
results2 <- filterCpGs(0.3, 0.2, 0.5, 3, 1)
visualizeResults(results2)
results3 <- filterCpGs(0.4, 0.2, 0.5, 2, 1)
visualizeResults(results3)
results4 <- filterCpGs(0.3, 0.2, 0.5, 2, 1)
visualizeResults(results4)
```



More tools

```{r}
createEmptyMatrix <- function(NUMCPGSITE, cpg_list, start_age, end_age) {
  B <- matrix(NA, nrow=NUMCPGSITE, ncol = end_age - start_age + 1)
  age_range <- start_age:end_age
  rownames(B) <- cpg_list
  column_names <- character(end_age - start_age + 1)
  for (i in 1:(end_age - start_age + 1)) {
    column_names[i] <- paste0("age_", i-1+start_age)
  }
  colnames(B) <- column_names
  return(B)
}

formatMatrix <- function(matrix_path) {
  temp <- readr::read_csv(matrix_path)
  #remove columns with all NA values
  temp <- temp[, colSums(is.na(temp)) != nrow(temp)]
  cpg_site_names <- temp[[1]]
  temp <- temp[, -1, with = FALSE]
  temp <- as.data.frame(temp)
  rownames(temp) <- cpg_site_names
  temp <- t(scale(t(temp)))
  temp <- as.matrix(temp)
  colnames(temp) <- gsub("age_", "", colnames(temp))
  return(temp)
}
```



# Make matrix of 1868 sites x total samples
```{r}
training_datasets <- c("GSE50660", "GSE90124", "GSE40279", "GSE67705", "GSE61256", "GSE73103", "GSE124366")
testing_datasets <- c("GSE114134", "GSE89253", "GSE53740", "GSE85568", "GSE106648", "GSE51057", "GSE124076")
all_datasets <- c(training_datasets, testing_datasets)
combined_path <- "/Users/kchen/OneDrive/Documents/methylation/combined18/"
# Read desired CpG sites from overlappingSites()
initial_cpg_list <- read.csv("7_3_overlap.csv") #check filter before running
initial_cpg_list <- initial_cpg_list$x
# Extract number of sites
NUMCPGSITE <- length(initial_cpg_list)
standardizeDataset <- function(df_name, desired_cpgs) {
  df <- readr::read_csv(paste0(combined_path, df_name, "_df.csv"))
  # Extract ages
  ages <- df[1, ]
  df <- df[-1, ]
  template <- data.table(cpg=desired_cpgs)
  # Reorder and filter by reference CpG sites
  df <- merge(template, df, by.x="cpg", by.y="cpg", all.x=TRUE, sort=FALSE)
  # Reattach ages
  df <- rbind(ages, df)
  
  return(df)
}

makeChunks <- function(datasets, output_folder) {
  temp <- data.frame()
  for (dataset in datasets) {
    df <- standardizeDataset(dataset, initial_cpg_list)
    if (ncol(temp) == 0) {
      temp <- df
    } else {
      temp <- cbind(temp, df)
    }
  }
  write.csv(temp, file = paste0(output_folder, "partition3_", 1, ".csv"), row.names = TRUE)
}

# Run code
makeChunks(all_datasets, "/Users/kchen/OneDrive/Documents/methylation/combined_split/")
```

# Reduce to a matrix of 1868 sites x 100 ages (avg betas)

```{r}
generateMatrix2 <- function(file_with_pre_filter, average, gender=FALSE, male=TRUE) {
  sites <- read.csv(file_with_pre_filter)
  sites <- sites[, -1]
  sites <- as.character(sites)
  #length: 288962
  NUMCPGSITE2 <- length(sites)
  #initialize output matrixs
  B <- matrix(NA, nrow=NUMCPGSITE2, ncol=101)
  age_range <- 0:100
  rownames(B) <- sites
  column_names <- character(101)
  for (i in age_range) {
    column_names[i+1] <- paste0("age_", i)
  }
  colnames(B) <- column_names
  temp <- readr::read_csv(paste0("/Users/kchen/OneDrive/Documents/methylation/combined_split/", "partition3_", 1, ".csv"))
  # Remove extra first column (automatic indices)
  temp <- temp[, -1]
  #filter for sites from correlation function
  des_rows <- which(temp$cpg...466 %in% sites)
  temp <- temp[c(1, des_rows), ] #keep ages in the first row
  # Extract CpG site list
  identifiers <- temp[-1, ncol(temp)]
  if (gender==TRUE) {
    if (male==TRUE) {
      temp <- temp[, colnames(temp) %in% male_samples]
    } else if (male==FALSE) {
      temp <- temp[, colnames(temp) %in% female_samples]
    }
  }
  for (j in 0:100) {
    # DEBUG // print(paste0("Split: ", i, " Age: ", j))
    des_cols <- which(temp[1, ]==j)
    # DEBUG // print(paste0("Columns at Age ", j, " : ", length(des_cols)))
    # Check if there are samples at age j
    if (length(des_cols)>0) {
      # Filter for age j
      age_data <- temp[, des_cols, drop = FALSE]
      # Drop age row
      age_data <- age_data[-1, ]
      
      #indice of identifiers corresponds to indice of age_data
      #age_data has samples as column names, no other identifiers
      # DEBUG // print(paste0("Dim of age_data: ", dim(age_data)))
      #this step takes the longest
      setDT(age_data)
      # Vectorized averaging
      if (average) {
        aggregated <- age_data[, .(final = rowMeans(.SD, na.rm = TRUE)), by = .I]
      }
      if (!average) {
        aggregated <- age_data[, .(final = apply(.SD, 1, sd, na.rm = TRUE)), by = .I]
      }        
      # DEBUG // print(paste0("Averaging done"))
      aggregated <- as.data.frame(aggregated)
      rownames(aggregated) <- as.character(identifiers$cpg...3952) #2441 for training, 1512 for testing
      # Reassignment to matrix B
      B[rownames(aggregated), paste0("age_", j)] <- aggregated$final
      # DEBUG //print(paste0("Split: ", i, ", Age: ", j, " updated"))
    }
  }
  return(B)
}

V <- generateMatrix2("7_3_overlap.csv", average=TRUE, gender=FALSE)
#write.csv(V, "7_3_small_matrix2.csv", row.names=TRUE)
M <- generateMatrix2("7_3_overlap.csv", average=TRUE, gender=TRUE, male=TRUE)
#write.csv(M, "7_3_male_matrix2.csv", row.names=TRUE)
Fe <- generateMatrix2("7_3_overlap.csv", average=TRUE, gender=TRUE, male=FALSE)
#write.csv(Fe, "7_3_female_matrix2.csv", row.names=TRUE)

#6_30_cpgfilter4.csv contains our filtered CpGs from the pipeline (file methylation_patterns.Rmd)
```

# Initialize Annotation Matrix
```{r}
B <- readr::read_csv("7_3_small_matrix2.csv")
B <- as.data.frame(B)
rownames(B) <- B[, 1]

clocks <- c("McEwen", "Wu", "Hannum", "Horvath", "Levine", "Belsky", "Teschendorff", "Shireby", "Lu")
clock_annotations <- data.frame(matrix(ncol = length(clocks), nrow = length(rownames(B))))
names(clock_annotations) <- clocks
rownames(clock_annotations) <- rownames(B)

clock_annotations$McEwen[rownames(B) %in% McEwen_CpGs] <- 1
clock_annotations$Wu[rownames(B) %in% Wu_CpGs] <- 1
clock_annotations$Hannum[rownames(B) %in% Hannum_CpGs] <- 1
clock_annotations$Horvath[rownames(B) %in% Horvath_CpGs] <- 1
clock_annotations$Levine[rownames(B) %in% Levine_CpGs] <- 1
clock_annotations$Belsky[rownames(B) %in% Belsky_CpGs] <- 1
clock_annotations$Teschendorff[rownames(B) %in% Teschendorff_CpGs] <- 1
clock_annotations$Shireby[rownames(B) %in% Shireby_CpGs] <- 1
clock_annotations$Lu[rownames(B) %in% Lu2_CpGs] <- 1

clock_annotations[is.na(clock_annotations)] <- 0

ha = rowAnnotation(df = clock_annotations, col = list(
    McEwen = c("0" = "gray", "1" = "purple"),
    Wu = c("0" = "gray", "1" = "orange"),
    Hannum = c("0" = "gray", "1" = "red"),
    Horvath = c("0" = "gray", "1" = "blue"),
    Levine = c("0" = "gray", "1" = "green"),
    Belsky = c("0" = "gray", "1" = "yellow"),
    Teschendorff = c("0" = "gray", "1" = "pink"),
    Shireby = c("0" = "gray", "1" = "magenta"),
    Lu = c("0" = "gray", "1" = "maroon")),
    simple_anno_size = unit(4, "mm"),
    show_legend=c(FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE, FALSE)
)
```


# Function to take in input matrix path, and number of clusters to plot
- returns matrix with cluster annotations in the last column
```{r}
mainPipeline <- function(matrix_path, num_clusters, row_title, plot=TRUE, save=FALSE, age_cutoff=100, clock=TRUE) {
  B <- formatMatrix(matrix_path)
  columns_to_keep <- as.numeric(colnames(B)) < age_cutoff
  B <- B[, columns_to_keep]
  #draw heatmap
  if (clock==TRUE) {
    ht <- Heatmap(B, 
        name = "Scaled Beta Values", 
        right_annotation = ha, 
        show_row_names = FALSE, 
        show_column_names = TRUE,
        cluster_rows = TRUE,
        cluster_columns = FALSE,
        row_title = row_title, 
        column_title = "Ages",
        column_names_gp = gpar(fontsize = 4, fontface="plain", fontfamily="serif"),
        column_names_rot = 0,
        row_split = num_clusters
  )
  } else {
    ht <- Heatmap(B, 
        name = "Scaled Beta Values", 
        #right_annotation = ha, 
        show_row_names = FALSE, 
        show_column_names = TRUE,
        cluster_rows = TRUE,
        cluster_columns = FALSE,
        row_title = row_title, 
        column_title = "Ages",
        column_names_gp = gpar(fontsize = 4, fontface="plain", fontfamily="serif"),
        column_names_rot = 0,
        row_split = num_clusters
  )
  }
  if (plot==TRUE) {
      if (save==TRUE) {
        png(paste0("july_figures/", matrix_path, num_clusters, "heatmap.png"))
      }
      ht <- draw(ht)
      if (save==TRUE) { 
        dev.off()
      }
  }
  cluster_assignments <- row_order(ht)
  cluster_vector <- unlist(lapply(1:length(cluster_assignments), function(i) {
    rep(i, length(cluster_assignments[[i]]))
  }))
  B_with_clusters <- cbind(B, Cluster=cluster_vector[order(unlist(cluster_assignments))])
  B_with_clusters <- as.data.frame(B_with_clusters)
  
  for (j in 1:num_clusters) {
    df <- B_with_clusters[which(B_with_clusters$Cluster==j),]
    long_df <- df %>%
      rownames_to_column(var="Sample") %>%
      pivot_longer(cols=-c("Sample"), names_to="Age", values_to="Beta") %>%
      mutate(Age=as.numeric(Age))
    long_df <- na.omit(long_df)
    
    fit <- lm(Beta ~ bs(Age, df=5), data=long_df)
    long_df$Fitted <- predict(fit)
    
    a <- ggplot(long_df, aes(x = Age, y = Beta)) +
      geom_point(size = 1) +
      geom_line(aes(y = Fitted), color = "blue") +
      #geom_smooth(method = "loess", formula = y ~ x, se = TRUE, color = "blue", fill = "lightblue", span = 0.75) +
      labs(title = paste("Cluster", j), x = "Age", y = "Scaled Beta Values") +
      theme_minimal() +
      ylim(-5, 5)
    if (plot==TRUE) {
      print(a)
    }  
    
    if (save==TRUE) {
      ggsave(filename=paste0("july_figures/", matrix_path, "Cluster_", j, ".png"), plot=a)
    }
  }
  return(B_with_clusters)
}
```

- "6_30_testing_matrix_v2.csv" is our selected CpGs
- "7_3_small_matrix2.csv" is combined Clock CpGs
- "7_3_male_matrix2.csv" is male Clock CpGs
- "7_3_female_matrix2.csv" is female Clock CpGs

```{r}
B_with_clusters <- mainPipeline("7_3_small_matrix2.csv", 4, "Clock CpGs", plot=TRUE, save=TRUE)
```
```{r}
B_with_clusters <- mainPipeline("7_3_small_matrix2.csv", 5, "Clock CpGs", plot=TRUE, save=FALSE)
B_with_clusters <- mainPipeline("7_3_small_matrix2.csv", 6, "Clock CpGs", plot=TRUE, save=FALSE)
```


```{r}
C_with_clusters <- mainPipeline("6_30_testing_matrix_v2.csv", 4, "Selected CpGs", plot=TRUE, save=TRUE, age_cutoff=100, clock=FALSE)
```

```{r}
M_with_clusters <- mainPipeline("7_3_male_matrix2.csv", 4, "Selected CpGs", plot=TRUE, save=FALSE, age_cutoff=80, clock=TRUE)
```

```{r}
Fe_with_clusters <- mainPipeline("7_3_female_matrix2.csv", 4, "Selected CpGs", plot=TRUE, save=FALSE, age_cutoff=80, clock=TRUE)
```


Clock CpG enrichment

```{r}
library(RColorBrewer)

#Fischer's exact test
fischers <- function(cont_table, clock_name) {
  p_values <- numeric(num_clusters)
  for (i in 1:num_clusters) {
    in_cluster <- cont_table[2*i-1, paste0(clock_name, "_clock_cpgs")]
    not_in_cluster <- cont_table[2*i, paste0(clock_name, "_clock_cpgs")]
    total_in_cluster <- sum(cont_table[2*i-1, -1])
    total_not_in_cluster <- sum(cont_table[2*i, -1])
    
    matrix_m <- matrix(c(in_cluster, total_in_cluster - in_cluster,
                        not_in_cluster, total_not_in_cluster - not_in_cluster),
                      nrow=2)
    
    fisher_test <- fisher.test(matrix_m)
    p_values[i] <- fisher_test$p.value
  }
  return(p_values)
}

clusterEnrichment <- function(matrix_with_clusters, num_clusters) {
  clusters <- lapply(1:num_clusters, function(i) rownames(matrix_with_clusters)[which(matrix_with_clusters[, ncol(matrix_with_clusters)] == i)])
  
  cont_table <- data.frame(
    clusters = character(),
    mcewen_clock_cpgs = integer(),
    wu_clock_cpgs = integer(),
    hannum_clock_cpgs = integer(),
    horvath_clock_cpgs = integer(),
    levine_clock_cpgs = integer(),
    belsky_clock_cpgs = integer(),
    teschendorff_clock_cpgs = integer(),
    shireby_clock_cpgs = integer(),
    lu_clock_cpgs = integer(),
    stringsAsFactors = FALSE
  )
  
  cpg_count_table <- matrix(0, nrow = num_clusters, ncol = length(clock_cpgs))
  colnames(cpg_count_table) <- names(clock_cpgs)
  rownames(cpg_count_table) <- paste("Cluster", 1:num_clusters)
  
  for (i in 1:num_clusters) {
    cluster_cpgs <- clusters[[i]]
    mcewen_clock_cpgs <- sum(cluster_cpgs %in% McEwen_CpGs)
    wu_clock_cpgs <- sum(cluster_cpgs %in% Wu_CpGs)
    hannum_clock_cpgs <- sum(cluster_cpgs %in% Hannum_CpGs)
    horvath_clock_cpgs <- sum(cluster_cpgs %in% Horvath_CpGs)
    levine_clock_cpgs <- sum(cluster_cpgs %in% Levine_CpGs)
    belsky_clock_cpgs <- sum(cluster_cpgs %in% Belsky_CpGs)
    teschendorff_clock_cpgs <- sum(cluster_cpgs %in% Teschendorff_CpGs)
    shireby_clock_cpgs <- sum(cluster_cpgs %in% Shireby_CpGs)
    lu_clock_cpgs <- sum(cluster_cpgs %in% Lu2_CpGs)
  
    cpg_count_table[i, ] <- c(mcewen_clock_cpgs, wu_clock_cpgs, hannum_clock_cpgs, horvath_clock_cpgs, levine_clock_cpgs, belsky_clock_cpgs, teschendorff_clock_cpgs, shireby_clock_cpgs, lu_clock_cpgs)

    cont_table <- rbind(cont_table, data.frame(
      Cluster = paste0("In Cluster ", i),
      mcewen_clock_cpgs = mcewen_clock_cpgs,
      wu_clock_cpgs = wu_clock_cpgs,
      hannum_clock_cpgs = hannum_clock_cpgs,
      horvath_clock_cpgs = horvath_clock_cpgs,
      levine_clock_cpgs = levine_clock_cpgs,
      belsky_clock_cpgs = belsky_clock_cpgs,
      teschendorff_clock_cpgs = teschendorff_clock_cpgs,
      shireby_clock_cpgs = shireby_clock_cpgs,
      lu_clock_cpgs = lu_clock_cpgs
    ))
    
    cont_table <- rbind(cont_table, data.frame(
      Cluster = paste0("Not In Cluster ", i),
      mcewen_clock_cpgs = length(McEwen_CpGs) - mcewen_clock_cpgs,
      wu_clock_cpgs = length(Wu_CpGs) - wu_clock_cpgs,
      hannum_clock_cpgs = length(Hannum_CpGs) - hannum_clock_cpgs,
      horvath_clock_cpgs = length(Horvath_CpGs) - horvath_clock_cpgs,
      levine_clock_cpgs = length(Levine_CpGs) - levine_clock_cpgs,
      belsky_clock_cpgs = length(Belsky_CpGs) - belsky_clock_cpgs,
      teschendorff_clock_cpgs = length(Teschendorff_CpGs) - teschendorff_clock_cpgs,
      shireby_clock_cpgs = length(Shireby_CpGs) - shireby_clock_cpgs,
      lu_clock_cpgs = length(Lu2_CpGs) - lu_clock_cpgs
    ))
  }
  
  p_values_mcewen <- fischers(cont_table, "mcewen")
  p_values_wu <- fischers(cont_table, "wu")
  p_values_hannum <- fischers(cont_table, "hannum")
  p_values_horvath <- fischers(cont_table, "horvath")
  p_values_levine <- fischers(cont_table, "levine")
  p_values_belsky <- fischers(cont_table, "belsky")
  p_values_teschendorff <- fischers(cont_table, "teschendorff")
  p_values_shireby <- fischers(cont_table, "shireby")
  p_values_lu <- fischers(cont_table, "lu")
  
  
  all_p_values <- c(p_values_mcewen, p_values_wu, p_values_hannum, p_values_horvath, p_values_levine, p_values_belsky, p_values_teschendorff, p_values_shireby, p_values_lu)
  
  #adjust for multiple comparisons
  adjusted_p_values <- p.adjust(all_p_values, method="BH")
  
  adjusted_p_values_mcewen <- adjusted_p_values[(3*num_clusters+1):(4*num_clusters)]
  adjusted_p_values_wu <- adjusted_p_values[(4*num_clusters+1):(5*num_clusters)]
  adjusted_p_values_hannum <- adjusted_p_values[1:num_clusters]
  adjusted_p_values_horvath <- adjusted_p_values[(num_clusters+1):(2*num_clusters)]
  adjusted_p_values_levine <- adjusted_p_values[(2*num_clusters+1):(3*num_clusters)]
  adjusted_p_values_belsky <- adjusted_p_values[(5*num_clusters+1):(6*num_clusters)]
  adjusted_p_values_teschendorff <- adjusted_p_values[(6*num_clusters+1):(7*num_clusters)]
  adjusted_p_values_shireby <- adjusted_p_values[(7*num_clusters+1):(8*num_clusters)]
  adjusted_p_values_lu <- adjusted_p_values[(8*num_clusters+1):(9*num_clusters)]
  
  
  heatmap_data <- data.frame(
    mcewen = adjusted_p_values_mcewen,
    wu = adjusted_p_values_wu,
    hannum = adjusted_p_values_hannum,
    horvath = adjusted_p_values_horvath,
    levine = adjusted_p_values_levine,
    belsky = adjusted_p_values_belsky,
    teschendorff = adjusted_p_values_teschendorff,
    shireby = adjusted_p_values_shireby,
    lu = adjusted_p_values_lu
  )
  rownames(heatmap_data) <- paste("Cluster", 1:num_clusters)
  
  annotation_matrix <- matrix("", nrow=num_clusters, ncol=9)
  colnames(annotation_matrix) <- colnames(heatmap_data)
  rownames(annotation_matrix) <- rownames(heatmap_data)
  
  for (i in 1:num_clusters) {
    for (j in 1:9) {
      if (heatmap_data[i, j] < 0.001) {
        annotation_matrix[i, j] <- "***"
      } else if (heatmap_data[i, j] < 0.01) {
        annotation_matrix[i, j] <- "**"
      } else if (heatmap_data[i, j] < 0.05) {
        annotation_matrix[i, j] <- "*"
      }
    }
  }
  
  p <- pheatmap(
    -log10(heatmap_data),
    cluster_rows = FALSE,
    cluster_cols = FALSE,
    display_numbers = annotation_matrix,
    number_color = "black",
    color = colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(100),
    main = "-log10 Adjusted P-Values for CpG Enrichment in Clusters",
  )
  
  print(p)
  
  cpg_count_table_df <- as.data.frame(cpg_count_table)
  colnames(cpg_count_table_df) <- names(clock_cpgs)
  rownames(cpg_count_table_df) <- paste("Cluster", 1:num_clusters)
  print(cpg_count_table_df)
}
```

```{r}
B_with_clusters <- mainPipeline("7_3_small_matrix2.csv", 4, "hello", plot=FALSE)
clusterEnrichment(B_with_clusters, 4)sum
```


Biological function enrichment

```{r}
annotation <- read.table("HM450.hg38.manifest.gencode.v36.tsv/HM450.hg38.manifest.gencode.v36.tsv", sep="\t", header=TRUE, quote="", comment.char="", fileEncoding = "UTF-8")
rownames(annotation) <- annotation$probeID

enrich <- function(sig_cpg, cluster) {
  temp_ann <- annotation[sig_cpg, ]
  genes <- temp_ann$genesUniq
  genes <- as.data.frame(genes)
  genes_unlist <- genes %>%
    separate_rows(genes, sep=";") %>%
    distinct()
  genes_unlist <- na.omit(genes_unlist)
  genes <- unlist(genes_unlist)
  
  #clean up gene version
  genes <- gsub("\\..*", "", genes)
  
  go_bp <- as.data.frame(enrichGO(gene = genes, OrgDb = org.Hs.eg.db, keyType = "SYMBOL", 
                      ont = "BP", pAdjustMethod = "BH", pvalueCutoff = 0.05))
  cluster_name <- paste0("Cluster ", cluster)
  if (nrow(go_bp)>0) {
    go_bp$Cluster <- cluster_name
  } 
  return(list(bp=go_bp))
}

GOEnrich <- function(matrix_with_clusters, num_clusters) {
  clusters <- lapply(1:num_clusters, function(i) rownames(matrix_with_clusters)[which(matrix_with_clusters[, ncol(matrix_with_clusters)] == i)])
  go_bp_results <- data.frame()
  for (i in 1:num_clusters) {
    cluster_name <- paste0("Cluster ", i)
    sig_cpg <- unlist(clusters[i])
    enrichment_results <- enrich(sig_cpg, i)
    
    go_bp_results <- rbind(go_bp_results, enrichment_results$bp)
  }
  return(go_bp_results)
}

go_bp_results <- GOEnrich(B_with_clusters, 4)
```

Plot Our Selected CpGs

```{r}
D <- readr::read_csv("6_30_testing_matrix_v2.csv")
D <- D[, colSums(is.na(D)) != nrow(D)]
cpg_site_names <- D[[1]]
D <- D[, -1, with = FALSE]
D <- as.data.frame(D)
rownames(D) <- cpg_site_names

overlap2 <- intersect(cpgs, rownames(D))
D <- D[!(rownames(D) %in% cpgs), ]

D <- t(scale(t(D)))
D <- as.matrix(D)
colnames(D) <- gsub("age_", "", colnames(D))

num_clusters = 4

ht <- Heatmap(D, 
        name = "Scaled Beta Values", 
        show_row_names = FALSE, 
        show_column_names = TRUE,
        cluster_rows = TRUE,
        cluster_columns = FALSE,
        row_title = "CpG Sites", 
        column_title = "Ages",
        column_names_gp = gpar(fontsize = 4, fontface="plain", fontfamily="serif"),
        column_names_rot = 0,
        row_split = num_clusters
)
ht <- draw(ht)

cluster_assignments <- row_order(ht)
cluster_vector <- unlist(lapply(1:length(cluster_assignments), function(i) {
  rep(i, length(cluster_assignments[[i]]))
}))

cluster_vector <- unlist(lapply(1:length(cluster_assignments), function(i) {
  rep(i, length(cluster_assignments[[i]]))
}))

D_with_clusters <- cbind(D, Cluster = cluster_vector[order(unlist(cluster_assignments))])
D_with_clusters <- as.data.frame(D_with_clusters)

for (j in 1:num_clusters) {
  df <- D_with_clusters[which(D_with_clusters$Cluster==j), ]
  long_df <- df %>%
    rownames_to_column(var="Sample") %>%
    pivot_longer(cols=-c("Sample"), names_to="Age", values_to="Beta") %>%
    mutate(Age=as.numeric(Age))

  long_df <- na.omit(long_df)
  fit <- lm(Beta ~ bs(Age, df = 5), data = long_df)
  long_df$Fitted <- predict(fit)

  a <- ggplot(long_df, aes(x = Age, y = Beta)) +
    geom_point(size = 1) +
    geom_line(aes(y = Fitted), color = "blue") +
    #geom_smooth(method = "loess", formula = y ~ x, se = TRUE, color = "blue", fill = "lightblue", span = 0.75) +
    labs(title = paste("Cluster", j), x = "Age", y = "Scaled Beta Values") +
    theme_minimal()
  print(a)
}
```
plot results for each gender
```{r}
D <- readr::read_csv("7_3_male_matrix.csv")
D <- D[, colSums(is.na(D)) != nrow(D)]
cpg_site_names <- D[[1]]
D <- D[, -1, with = FALSE]
D <- as.data.frame(D)
rownames(D) <- cpg_site_names

D <- t(scale(t(D)))
D <- as.matrix(D)
colnames(D) <- gsub("age_", "", colnames(D))

num_clusters = 4

ht <- Heatmap(D, 
        name = "Scaled Beta Values", 
        show_row_names = FALSE, 
        show_column_names = TRUE,
        cluster_rows = TRUE,
        cluster_columns = FALSE,
        row_title = "CpG Sites", 
        column_title = "Ages",
        column_names_gp = gpar(fontsize = 4, fontface="plain", fontfamily="serif"),
        column_names_rot = 0,
        row_split = num_clusters
)
ht <- draw(ht)

cluster_assignments <- row_order(ht)
cluster_vector <- unlist(lapply(1:length(cluster_assignments), function(i) {
  rep(i, length(cluster_assignments[[i]]))
}))

cluster_vector <- unlist(lapply(1:length(cluster_assignments), function(i) {
  rep(i, length(cluster_assignments[[i]]))
}))

D_with_clusters <- cbind(D, Cluster = cluster_vector[order(unlist(cluster_assignments))])
D_with_clusters <- as.data.frame(D_with_clusters)

for (j in 1:num_clusters) {
  df <- D_with_clusters[which(D_with_clusters$Cluster==j), ]
  long_df <- df %>%
    rownames_to_column(var="Sample") %>%
    pivot_longer(cols=-c("Sample"), names_to="Age", values_to="Beta") %>%
    mutate(Age=as.numeric(Age))

  long_df <- na.omit(long_df)
  fit <- lm(Beta ~ bs(Age, df = 5), data = long_df)
  long_df$Fitted <- predict(fit)

  a <- ggplot(long_df, aes(x = Age, y = Beta)) +
    geom_point(size = 1) +
    geom_line(aes(y = Fitted), color = "blue") +
    #geom_smooth(method = "loess", formula = y ~ x, se = TRUE, color = "blue", fill = "lightblue", span = 0.75) +
    labs(title = paste("Cluster", j), x = "Age", y = "Scaled Beta Values") +
    theme_minimal()
  print(a)
}
```

```{r}
D <- readr::read_csv("7_3_female_matrix.csv")
D <- D[, colSums(is.na(D)) != nrow(D)]
cpg_site_names <- D[[1]]
D <- D[, -1, with = FALSE]
D <- as.data.frame(D)
rownames(D) <- cpg_site_names

D <- t(scale(t(D)))
D <- as.matrix(D)
colnames(D) <- gsub("age_", "", colnames(D))

num_clusters = 4

ht <- Heatmap(D, 
        name = "Scaled Beta Values", 
        show_row_names = FALSE, 
        show_column_names = TRUE,
        cluster_rows = TRUE,
        cluster_columns = FALSE,
        row_title = "CpG Sites", 
        column_title = "Ages",
        column_names_gp = gpar(fontsize = 4, fontface="plain", fontfamily="serif"),
        column_names_rot = 0,
        row_split = num_clusters
)
ht <- draw(ht)

cluster_assignments <- row_order(ht)
cluster_vector <- unlist(lapply(1:length(cluster_assignments), function(i) {
  rep(i, length(cluster_assignments[[i]]))
}))

cluster_vector <- unlist(lapply(1:length(cluster_assignments), function(i) {
  rep(i, length(cluster_assignments[[i]]))
}))

D_with_clusters <- cbind(D, Cluster = cluster_vector[order(unlist(cluster_assignments))])
D_with_clusters <- as.data.frame(D_with_clusters)

for (j in 1:num_clusters) {
  df <- D_with_clusters[which(D_with_clusters$Cluster==j), ]
  long_df <- df %>%
    rownames_to_column(var="Sample") %>%
    pivot_longer(cols=-c("Sample"), names_to="Age", values_to="Beta") %>%
    mutate(Age=as.numeric(Age))

  long_df <- na.omit(long_df)
  fit <- lm(Beta ~ bs(Age, df = 5), data = long_df)
  long_df$Fitted <- predict(fit)

  a <- ggplot(long_df, aes(x = Age, y = Beta)) +
    geom_point(size = 1) +
    geom_line(aes(y = Fitted), color = "blue") +
    #geom_smooth(method = "loess", formula = y ~ x, se = TRUE, color = "blue", fill = "lightblue", span = 0.75) +
    labs(title = paste("Cluster", j), x = "Age", y = "Scaled Beta Values") +
    theme_minimal()
  print(a)
}
```



BELOW IS OLD CODE

Reformat new matrix
- Remove ages 0, 1, 2 because values are very sparse
- use linear interpolation for the rest of them to preserve patterns (removing all NAs reduces the list of CpGs down to 267 out of 1074)
- also, for some sites missing values at ages, 87-96, so filled in with last value (only for clustering)
- Heatmap is plotted without any interpolation
```{r}
"""
library(zoo)
B <- readr::read_csv("6_27_small_matrix.csv")
B <- B[, colSums(is.na(B)) != nrow(B)]
cpg_site_names <- B[[1]]
B <- B[, -1, with = FALSE]
B <- as.data.frame(B)
rownames(B) <- cpg_site_names
B <- B[, -(1:3)]
B_interpolated <- data.frame(t(apply(B, 1, function(x) {
    interpolated <- na.approx(x, na.rm = FALSE) 
    first_non_na <- which(!is.na(interpolated))[1]
    if (!is.na(first_non_na)) {
        interpolated[1:first_non_na] <- interpolated[first_non_na]
    }
    last_non_na <- which(!is.na(interpolated))[length(which(!is.na(interpolated)))]
    if (!is.na(last_non_na)) {
        interpolated[last_non_na:length(interpolated)] <- interpolated[last_non_na]
    }
    return(interpolated)
})))
colnames(B_interpolated) <- colnames(B)
#scale
scaled_B <- t(scale(t(B_interpolated)))
df <- as.data.frame(scaled_B)
"""
```

Hierarchical Clustering

```{r}
"""
d <- dist(df, method = "euclidean")  # Compute the distance matrix
hc <- hclust(d, method = "ward.D2")  # Perform hierarchical clustering

# Plot the dendrogram
plot(hc, main = "Dendrogram", xlab = "Sample index", ylab = "Height")
"""
```
```{r}
"""
clusters <- cutree(hc, k = 4)
df$Cluster <- as.factor(clusters)
df <- df[order(df$Cluster), ]
row_order <- rownames(df)

scaledB <- t(scale(t(B)))
df2 <- as.data.frame(scaledB)
original <- df2[row_order, ]
"""
```

Simple heatmap

```{r}
"""
library(pheatmap)
pheatmap(as.matrix(original),  # Exclude cluster column
         scale = "none",
         show_rownames = FALSE,
         show_colnames = TRUE,
         cluster_cols = FALSE,
         cluster_rows = FALSE,
         color = colorRampPalette(c("blue", "white", "red"))(100),
         fontsize_col = 6,
         angle_col = 45,
         main="Clock CpGs",
         breaks = seq(-5, 5, length.out = 101)
)
"""
```

Read annotations

```{r}
```




